Notes on robust solvers for systems of coupled multiscale PDEs

### 2019-08-02 solver/problem_gamma.tex
update

Let $V=H^1{(\Gamma)}$ We consider the minimization problem
\[
\min_{u\in V}\quad\norm{u}^2_1 - \int_{\Gamma}f u\,\mathrm{d}x\quad\mbox{subject to}\quad
u-g\mbox{ in }\Gamma.
\]
Denoting $p$ the Lagrange multiplier for the constraint the extremal points
of the related Lagrangian are found as a solution to
%
\begin{equation}\label{eq:A1}
\begin{pmatrix}f\\g\end{pmatrix}=
\mathcal{A}_1\begin{pmatrix}u\\p\end{pmatrix}=
%
\begin{pmatrix}
  -\Delta + I & I\\
  I           &  \\
\end{pmatrix}
\begin{pmatrix}u\\p\end{pmatrix}.
\end{equation}
%
Setting $Q=H^{-1}(\Gamma)$, the operator $\mathcal{A}_1:W\rightarrow W^{\prime}$
can be seen to be an isomorphism with $W=V\times Q$. In particular the inf-sup
condition holds with constant 1 (by definition of the $H^{-1}$ norm).

@@

Example \ref{ex:stokes} motivates the need to find a stabilizing term
for the P1-P0 discretization of $\mathcal{A}_1$. In addition, the stabilization
should be consisten so that to not ruin the approximation properties of the
formulation.

>>>>

Let $V=H^1{(\Gamma)}$ We consider the minimization problem
\[
\min_{u\in V}\quad\norm{u}^2_1 - \int_{\Gamma}f u\,\mathrm{d}x\quad\mbox{subject to}\quad
u=g\mbox{ in }\Gamma.
\]
Denoting $p$ the Lagrange multiplier for the constraint the extremal points
of the related Lagrangian are found as a solution to
%
\begin{equation}\label{eq:A1}
\begin{pmatrix}f\\g\end{pmatrix}=
\mathcal{A}_1\begin{pmatrix}u\\p\end{pmatrix}=
%
\begin{pmatrix}
  -\Delta + I & I\\
  I           &  \\
\end{pmatrix}
\begin{pmatrix}u\\p\end{pmatrix}.
\end{equation}
%
Setting $Q=H^{-1}(\Gamma)$, the operator $\mathcal{A}_1:W\rightarrow W^{\prime}$
can be seen to be an isomorphism with $W=V\times Q$. In particular the inf-sup
condition holds with constant 1 (by definition of the $H^{-1}$ norm).

@@

Example \ref{ex:stokes} motivates the need to find a stabilizing term
for the P1-P0 discretization of $\mathcal{A}_1$. In addition, the stabilization
should be consistent so as to not ruin the approximation properties of the
formulation.

\subsection{Stabilized $\mathcal{A}_1$} We construct the stabilizing term
based on two observations. (i) In \S\ref{sec:problem_omega} the stabilization
for $p\in Q_h\subset H^{-1/2}$ in takes the form the facet integral with integrand $h^{2}\jump{p}\jump{q}$.
(ii) For P0 element $-\Delta_h$ is spectrally equivalent to operator involving
facet integrals with integrand $h^{-1}\jump{p}\jump{q}$. In turn we assume
that the stabilization for $H^s$ norms takes the form $h^{p}\jump{p}\jump{q}$
and by interpolation from (i) and (ii) $p=-2s+1$. Thus the stabilizing term for $H^{-1}$ shall
involve the cube of the mesh size and we consider a discrete formulation of \eqref{eq:A1}
in terms of the operator
%
\[
\langle\mathcal{A}_{1, h}\begin{pmatrix}u\\p\end{pmatrix},
  \begin{pmatrix}v\\q\end{pmatrix}
    \rangle
    =
    \int_{\Gamma}\nabla u\cdot\nabla v+uv\,\mathrm{d}x+\int_{\Gamma}uq + vp\,\mathrm{d}x-
    \sum_{F\in\mathcal{F}}h^3\jump{p}\jump{q}\,\mathrm{d}s.
\]
The preconditioner for $\mathcal{A}_{1, h}$ is then the Riesz mapping which
includes the negative stabilized term
\[
\langle\mathcal{B}^{-1}_{1, h}\begin{pmatrix}u\\p\end{pmatrix},
  \begin{pmatrix}v\\q\end{pmatrix}
    \rangle
    =
    \int_{\Gamma}\nabla u\cdot\nabla v+uv\,\mathrm{d}x+
    \langle(-\Delta+I)^{-1}p, q\rangle+
    \sum_{F\in\mathcal{F}}h^3\jump{p}\jump{q}\,\mathrm{d}s.
\]

Table \ref{tab:A1} shows error convergence of the proposed formulation
together with the condition number of the preconditioner problem. We remark
that the manufactured solution on the unit interval has $g=0$ so that $p=f=\cos(\pi x)$
and $u=0$ is the solution. It can be seen that the formulation is stable and
yields converging solutions.

\begin{table}
  \footnotesize{
    \begin{center}
      \begin{tabular}{c|cc|c}
        \hline
        $h$ & $\norm{u-u_h}_1$ & $\norm{p-p_h}_0$ & $\kappa$ \\
        \hline
% 2.50E-01& 6.4083E-02(--)  & 2.4243E-01(--)  & 2.65607  \\
% 1.25E-01& 5.5055E-03(3.54)& 8.0279E-02(1.59)& 2.66149  \\
% 6.25E-02& 3.5103E-04(3.97)& 4.0044E-02(1.00)& 2.66364  \\
% 3.12E-02& 2.1949E-05(4.00)& 2.0040E-02(1.00)& 2.66582  \\
 1.56E-02& 1.3715E-06(4.00)& 1.0020E-02(1.00)& 2.66592  \\
 7.81E-03& 8.5716E-08(4.00)& 5.0100E-03(1.00)& 2.66596  \\
 3.91E-03& 5.3572E-09(4.00)& 2.5050E-03(1.00)& 2.66599  \\
 1.95E-03& 3.3482E-10(4.00)& 1.2525E-03(1.00)& 2.66599  \\
 9.77E-04& 2.0926E-11(4.00)& 6.2625E-04(1.00)& 2.66599  \\
 4.88E-04& 1.3079E-12(4.00)& 3.1312E-04(1.00)& 2.66575  \\
        \hline
      \end{tabular}
    \end{center}
  }
  \caption{Conditioning and approximation properties of stabilized
    formulation of \eqref{eq:A1} using P1-P0 elements.
  }
  \label{tab:A1}
\end{table}
